### 2017-04-11 src/stats6.tex
Created SS6 chapter structure.

\chapter*{Statistics 6}

\newpage
\section{Experimental Design}

\newpage
\section{Analysis of Variance}

\newpage
\section{Statistical Process Control}

\newpage
\section{Acceptance Sampling}