### 2017-05-31 src/mstats3.tex
Removed new pages and created files for MS03 and MS04.

\cchapter{Maths Statistics 3}

\newpage
\section{Geometric and Exponential Distributions}