### 2017-05-31 src/mstats3.tex
Added section titles.

\cchapter{Maths Statistics 3}

\section{Geometric and Exponential Distributions}
\section{Estimators}
\section{Estimation}
\section{Hypothesis Testing}
\section{Chi-Squared ($\chi^2$) Goodness of Fit Testing}