### 2017-05-31 src/mstats3.tex
Added geometic distributions section.

\cchapter{Maths Statistics 3}

\section{Geometric and Exponential Distributions}

    \subsection{Geometric Distributions}
        Geometric distribution is like a binomial distribution however, instead of counting the number of sucesses, it counts the number of trials until a success is obtained.
        \\\\
        Conditions for a Geometric Distribution:
        \begin{itemize}
        \item Only two possible outcomes (success or failure).
        \item Trials are independent.
        \item Probabilty of success is the same for each trial.
        \end{itemize}
        The probability density function for a Geometric Distribution is as follows:
        \begin{align*}
        P(x) &= p(1-p)^{x-1}
        \end{align*}
        Which makes sence since the probability of a success is the $x$ times the probability of failure times the single probability of success.
        \\\\
        The mean of a Geometric Distribution is dirivied as follows.
        \begin{align*}
        \text{Let }q &= 1-p                     \\
        E(X) &= \sum{x_i p_i}                   \\
             &= p + 2pq + 3pq^2 + 4pq^3 + ...   \\
             &= p(1 + 2q + 3q^2 + 4q^3 + ...)   \\
             &= p(1-q)^{-2}                     \\
             &= \dfrac{p}{p^2}                  \\
             &= \dfrac{1}{p}
        \end{align*}
        \\\\
        And the varience can be found using $E(X^2) - E(X)^2 = Var(X)$
        \begin{align*}
        E(X^2) &= \dfrac{2-p}{p^2} \text{  (using the method above)}    \\
        Var(X) &= \dfrac{2-p}{p^2} - \left(\dfrac{1}{p}\right)^2        \\
               &= \dfrac{2-p-1}{p^2}                                    \\
               &= \dfrac{1-p}{p^2}                                      \\
        \end{align*}
    \subsection{Exponential Distributions}
        (See Exponential Distribution section)

\section{Estimators}
\section{Estimation}
\section{Hypothesis Testing}
\section{Chi-Squared ($\chi^2$) Goodness of Fit Testing}